---
title: "Trees in R"
output: pdf_document
author: "Leann M. Biancani"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Install required packages (if needed)
#install.packages("BiocManager")
#library("BiocManager")
#BiocManager::install('ggtree') # could not install on URI Rstudio server
#BiocManager::install('treeio')
#install.packages('ggplot2')


## load necessary packages:
library('treeio')
library('ggtree')
library('ggplot2')

```


```{r tree figs, echo=FALSE, warning=FALSE}

# import tree file

pi_m20_nogap <- "TreeFiles/RAxML_bipartitionsBranchLabels.tree_pi_m20_nogap"
tree <- read.raxml(pi_m20_nogap)
tree <- root(tree, node = 66, resolve.root = TRUE) # reroot tree at outgroup
#is.rooted(tree) # check if tree is rooted

#ggtree(tree)

# start plotting tree:
ggtree(tree)  +
  xlim(0,2) + #adjust plot area to allow more space for labels
  geom_treescale() + # add a scalebar
  geom_tiplab(size =3) + # add tip labels
  geom_rootedge(rootedge = 0.1) + # display the root edge
  geom_text(aes(label=bootstrap), size = 2, hjust=-.1) + # add bootstrap support to nodes (shifting labels slightly to the right)
  #geom_text(aes(label=node), hjust=1) + # prints node numbers (info for rerooting: marsupial clade = node 66)
  # Add clade labels
  geom_cladelabel(node=55, label="Boreoeutheria", color="#08A69E", hjust=-1.25) +
  geom_cladelabel(node=49, label="Xenarthra", color="#0276BA", hjust=-2.5) +
  geom_cladelabel(node=43, label="Afrotheria", color="#1EB101", hjust=-1.75)

## Notes:
## Boreoeutheria: color = "#08A69E" node = 55
## Xenarthra: color = "#0276BA"  node = 49
## Afrotheria: color = "#1EB101" node = 43

```

Figure 2. Maximum Likelihood Phylogenetic Analysis performed in RAxML for 30 placental mammal and 10 marsupial outgroup taxa. Orthologous loci were obtained from raw genomic reads using the SISRS pipeline. Support values represent 100 bootstrap replicates. Placental mammal clades are indicated by color-coded vertical lines.  These preliminary results appear to support the Afrotheria hypothesis for the branching pattern at the root of placental mammals. 